Retrofitting Effect Handlers onto OCaml

### 2020-11-06 retro-concurrency.tex
Add minor comment

The results are presented in Table~\ref{tab:micro_noperform}. They show that
using effect handlers (concurrency monad) is 13.77$\times$ (67.09$\times$) more
expensive than the idiomatic implementation using non-tail calls. Concurrency
monad suffers due to the heap allocation of continuation frames (which need to
be GCed), where as effect handlers benefit from stack allocation of the frames.
We note that a compiler that uses CPS IR will be faster than the concurrency
monad implementation due to optimisations to reduce the heap allocation of
continuation frames. But Farvardin et al.~\cite{} show that CPS with
optimisations is still slower than using the call stack.

@@

Next we look at benchmarks that utilize non-local control flow. First, we
quantify the overhead individual operations in effect handling. Consider the
following annotated code:
\begin{lstlisting}
effect E : unit
(* a *) match (* b *) perform E (* d *) with
| v -> (* e *) v
| effect E k -> (* c *) continue k ()
\end{lstlisting}
The sequence |a-b| involves allocating a new fiber and switching to it. |b-c|
is performing the effect and handling it. |c-d| is resuming the continuation.
|d-e| is returning from the fiber with a value and freeing the fiber. We
measured the time taken to execute these sequences using |perf| support for
cycle-accurate tracing on modern Intel processors. We executed 10 iterations of
the code, ignoring 3 warm-up runs. We observed that the sequences
|a-b|, |b-c|, |c-d| and |d-e| took 48ns, 10ns, 13ns and 19ns, respectively. For
reference, the idle memory latency for the local NUMA node is 93.2ns.

>>>>

The results are presented in Table~\ref{tab:micro_noperform}. They show that
using effect handlers (concurrency monad) is 13.77$\times$ (67.09$\times$) more
expensive than the idiomatic implementation using non-tail calls. Concurrency
monad suffers due to the heap allocation of continuation frames (which need to
be GCed), where as effect handlers benefit from stack allocation of the frames.
Our concurrency monad, and indeed monadic concurrency libraries such as Lwt and
Async, also has other downsides; due to the lack of stack, exceptions,
backtraces, and DWARF unwinding are no longer useful.

We note that a compiler that uses CPS IR will be faster than the concurrency
monad implementation due to optimisations to reduce the heap allocation of
continuation frames. But Farvardin et al.~\cite{} show that CPS with
optimisations is still slower than using the call stack.

@@

Next we look at benchmarks that utilize non-local control flow. First, we
quantify the overhead individual operations in effect handling. Consider the
following annotated code:
\begin{lstlisting}
effect E : unit
(* a *) match (* b *) perform E (* d *) with
| v -> (* e *) v
| effect E k -> (* c *) continue k ()
\end{lstlisting}
The sequence \texttt{a-b} involves allocating a new fiber and switching to it.
\texttt{b-c} is performing the effect and handling it. \texttt{c-d} is resuming
the continuation. \texttt{d-e} is returning from the fiber with a value and
freeing the fiber. We measured the time taken to execute these sequences using
|perf| support for cycle-accurate tracing on modern Intel processors. We
executed 10 iterations of the code, ignoring 3 warm-up runs. We observed that
the sequences \texttt{a-b}, \texttt{b-c}, \texttt{c-d} and \texttt{d-e} took
48ns, 10ns, 13ns and 19ns, respectively. For reference, the idle memory latency
for the local NUMA node is 93.2ns.
